---
title: "Murrelet & Raven Detection Models"
author: "Tam Ta & Amelia J. DuVall"
date: "4/13/2022"
output: pdf_document
---

# TODO
# add site effect?
# add distance from acoustic unit to camera for acoustic models? 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load libraries
library(tidyverse)
library(here)
library(lme4)

## load data
camera_full <- read.csv(here("data", "camera_dat_all_images.csv")) 
camera_set <- read.csv(here("data", "camera_dat_image_set.csv")) 
camera_hourly <- read.csv(here("data", "camera_hourly_det.csv"))
acoustic <- read.csv(here("data", "acoustic", "acoustic_cam_tod.csv")) 
```
# Camera Trap Data 

## Model #1: SCMU Camera Detections
Binomial generalized linear mixed model to understand effect of image type (motion-capture versus timelapse), time of day (dawn, day, dusk) and time of year (julian day, quadratic effect) plus random effect of camera id (n = 7) on ability to detect murrelets. 

```{r mod1}
# test <- camera_hourly %>%
#   mutate(tod = ifelse(tod == "day", "aday", tod))
# m1 <- glmer(SCMU_hourly_det ~ image_type + tod + jday + jday^2 + (1 | cam_id), 
#             data = test, family = binomial)

m1 <- glmer(SCMU_hourly_det ~ image_type + tod + jday + jday^2 + (1 | cam_id), 
            data = camera_hourly, family = binomial)
summary(m1)

## day of year as a quadratic
# plot(camera_hourly$jday, camera_hourly$jday) 
# plot(camera_hourly$jday, camera_hourly$jday + camera_hourly$jday^2)
# plot(camera_hourly$jday, (-0.14*camera_hourly$jday +  ((-0.14^2)*camera_hourly$jday^2)))
```

## Model #2: CORA Camera Detections
Binomial generalized linear mixed model to understand effect of image type (motion-capture versus timelapse), time of day (dawn, day, dusk) and time of year (julian day, quadratic effect) plus random effect of camera id (n = 7) on ability to detect ravens.
```{r mod2}
m2 <- glmer(CORA_hourly_det ~ image_type + tod + jday + jday^2 + (1 | cam_id), 
            data = camera_hourly, family = binomial)
summary(m2)

```

## Model #3: Total images
Poisson (or negative binomial?) generalized linear mixed model to understand effect of total number of images on ability to detect murrelets (repeat for ravens?). 

Plot the data, suspect will be highly overdispersed, thus negative binomial): glmmb (mixed effect package that handles negative binomial)
```{r m3}
plot()


```


## Model #4: Image sets
Poisson generalized linear mixed model to understand effect of number of image sets on ability to detect murrelets (repeat for ravens?). 
```{r m4}

```

# Acoustic Data

## Model #5: SCMU Acoustic Detections
Binomial generalized linear mixed model to understand effect of time since camera detection, before/after/during camera detection, time of day (dawn, day, dusk), time of year (julian day, quadratic effect), distance from acoustic unit to camera (?) plus random effect of camera id (n = 7) on ability to detect murrelets on acoustic units given known camera detection.

```{r m5}

```

## Model #6: CORA Acoustic Detections
Binomial generalized linear mixed model to understand effect of time since camera detection, before/after/during camera detection, time of day (dawn, day, dusk), time of year (julian day, quadratic effect), distance from acoustic unit to camera (?) plus random effect of camera id (n = 7) on ability to detect ravens on acoustic units given known camera detection.
```{r m6}

```