---
title: "Murrelet & Raven Detection Models"
author: "Tam Ta & Amelia J. DuVall"
date: "4/13/2022"
output: pdf_document
---

# TODO
# add site effect?
# add distance from acoustic unit to camera for acoustic models? 
# figure out camera ID for acoustic detections
# add categorical column: before = time.since2 is pos or after = time.sinc2 is neg
# think camera before acoustic
# add 0 for motion
# add log(photo count)
# filter for timelapse for acoustic models
# table of deployment dates


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load libraries
library(tidyverse)
library(here)
library(lme4)

## load data
camera_hourly <- read.csv(here("data", "camera_hourly_mod.csv"))
acoustic <- read.csv(here("data", "acoustic_cam_tod.csv")) 

#camera_hourly[intersect(which(camera_hourly$image_type == "timelapse"),which(camera_hourly$photo_count < 6))   ]

# table(camera_hourly$photo_count, camera_hourly$image_type)
chk <- camera_hourly %>%
  filter(image_type == "timelapse" & photo_count != 6)
```
# Camera Trap Data 

## Model #1: SCMU Camera Detections
Binomial generalized linear mixed model to understand effect of image type (motion-capture versus timelapse), time of day (dawn, day, dusk) and time of year (julian day, quadratic effect) plus random effect of camera id (n = 7) on ability to detect murrelets within one-hour time periods. 

```{r mod1}
# test <- camera_hourly %>%
#   mutate(tod = ifelse(tod == "day", "aday", tod))
# m1 <- glmer(SCMU_hourly_det ~ image_type + tod + jday + jday^2 + (1 | cam_id), 
#             data = test, family = binomial)
#image_type + photo_count %in% image_type
#image_type + image_type : photo_count
#image_type + image_type : I(log(photo_count)) 
#I have standardized jday so that we don't have numerical problems when running jday^2 

camera_hourly$jday.stand  <- (camera_hourly$jday - mean(camera_hourly$jday))/sd(camera_hourly$jday)


m1 <- glmer(SCMU_hourly_det ~ image_type + tod + jday.stand + I(jday.stand^2) + 
              (1 | cam_id), 
            data = camera_hourly, family = binomial)
summary(m1)
confint(m1,level=0.95)

m1.5 <- glmer(SCMU_hourly_det ~ image_type + image_type : photocount + tod + jday.stand + I(jday.stand^2) +
              (1 | cam_id), 
            data = camera_hourly, family = binomial)
summary(m1.5)

## day of year as a quadratic
# plot(camera_hourly$jday, camera_hourly$jday) 
# plot(camera_hourly$jday, camera_hourly$jday + camera_hourly$jday^2)
# plot(camera_hourly$jday, (-0.14*camera_hourly$jday +  ((-0.14^2)*camera_hourly$jday^2)))
```

## Model #2: CORA Camera Detections
Binomial generalized linear mixed model to understand effect of image type (motion-capture versus timelapse), time of day (dawn, day, dusk) and time of year (julian day, quadratic effect) plus random effect of camera id (n = 7) on ability to detect ravens within one-hour time periods. 
```{r mod2}
m2 <- glmer(CORA_hourly_det ~ image_type + image_type:photo_count + tod + jday.stand + I(jday.stand^2) + (1 | cam_id), 
            data = camera_hourly, family = binomial)
m2.5 <- glmer(CORA_hourly_det ~ image_type + tod + jday.stand + I(jday.stand^2) + (1 | cam_id), 
            data = camera_hourly, family = binomial)
summary(m2.5)

```

<!-- ## Model #3: Total images -->
<!-- Poisson (or negative binomial?) generalized linear mixed model to understand effect of image type (motion-capture versus timelapse), time of day (dawn, day, dusk) and time of year (julian day, quadratic effect) plus random effect of camera id (n = 7) on ability to detect murrelets using all images.  -->

<!-- Plot the data, suspect will be highly overdispersed, thus negative binomial): glmmb (mixed effect package that handles negative binomial) -->

<!-- Repeat for ravens? Need to group detections somehow? -->
<!-- ```{r m3} -->
<!-- # plot(camera_full$SCMU) -->

<!-- m3 <- glmer(SCMU ~ image_type + tod + jday + jday^2 + (1 | cam_id),  -->
<!--             data = camera_full, family = binomial) -->
<!-- summary(m3) -->


<!-- ``` -->


<!-- ## Model #4: Image sets -->
<!-- Poisson generalized linear mixed model to understand effect of image type (motion-capture versus timelapse), time of day (dawn, day, dusk) and time of year (julian day, quadratic effect) plus random effect of camera id (n = 7) on ability to detect murrelets using image sets within 10-minute detection windows.  -->

<!-- ```{r m4} -->
<!-- m4 <- glmer(SCMU ~ image_type + tod + jday + jday^2 + (1 | cam_id),  -->
<!--             data = camera_set, family = binomial) -->
<!-- summary(m4) -->
<!-- ``` -->

# Acoustic Data

## Model #5: SCMU Acoustic Detections
Binomial generalized linear mixed model to understand effect of time since camera detection, before/after/during camera detection, time of day (dawn, day, dusk), time of year (julian day, quadratic effect), distance from acoustic unit to camera (?) plus random effect of camera id (n = 7) on ability to detect murrelets on acoustic units given known camera detection.

```{r m5}
acoustic <- acoustic %>% mutate(time_int = ifelse(time_int == "during", "aduring", time_int))

#I have standardized jday so that we don't have numerical problems when running jday^2 
acoustic$jday.stand  <- (acoustic$jday - mean(acoustic$jday)) / sd(acoustic$jday)

m5 <- glmer(SCMU ~ time_int + tod + jday.stand + I(jday.stand^2) + #distance + 
              (1 | cam_id), 
            data = acoustic, family = binomial)
summary(m5)


```

## Model #6: CORA Acoustic Detections
Binomial generalized linear mixed model to understand effect of time since camera detection, before/after/during camera detection, time of day (dawn, day, dusk), time of year (julian day, quadratic effect), distance from acoustic unit to camera (?) plus random effect of camera id (n = 7) on ability to detect ravens on acoustic units given known camera detection.
```{r m6}
cora <- acoustic %>% filter(CORA == 1) # there are only 26 CORA acoustic detections!

m6 <- glmer(CORA ~ time_int + tod + jday.stand + I(jday.stand^2) + #distance +
               (1 | cam_id),
             data = acoustic, family = binomial)
 summary(m6)
```